---
title: "MAMMALPITCH"
author: "Saanie Zahra Naqvi"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Slide 1 — Introduction
The **Mammal Body Mass Predictor** is an interactive R Shiny app that estimates a mammal’s body weight using **brain weight** and **sleep duration**.  
It combines biological reasoning with data science principles.

---

## Slide 2 — Dataset and Source
Data come from the `msleep` dataset in **ggplot2**, originally published by  
Dr. V. M. Savage et al. in *Nature* (2008).  

It contains physiological traits for ~80 mammal species:
- Body Weight (kg)
- Brain Weight (g)
- Sleep Duration (hours/day)
- Feeding Type

```{r}
summary(ggplot2::msleep[, c("bodywt", "brainwt", "sleep_total")])
```

---

## Slide 3 — Model and Logic
- User inputs brain weight and sleep duration  
- The server fits two regression models:
  - Linear  
  - Log-transformed (for non-linear scaling)
- Output: estimated body weight  
- Visualization: scatter plot + highlighted input point

---

## Slide 4 — App Preview
Visit the live app:  
[https://saanienaqvi.shinyapps.io/mammal_mass_predictor/](https://saanienaqvi.shinyapps.io/mammal_mass_predictor/)

```{r echo=FALSE}
library(ggplot2)
data(msleep)
plot(msleep$brainwt, msleep$bodywt, log="xy", col="skyblue", pch=19,
     xlab="Brain Weight (g)", ylab="Body Weight (kg)",
     main="Brain vs Body Weight in Mammals")
```

---

## Slide 5 — Conclusion
🧠 *Biology meets Data Science*  
- Demonstrates how physiology and behavior connect  
- Real dataset, reproducible analysis  
- Built using **R Shiny** and **ggplot2**  
- Deployed on **shinyapps.io** and presented on **RPubs**
